---
title: "Week 4 In-class Exercises"
output:
  html_document:
    df_print: paged
---

```{r setup}
knitr::opts_knit$set(root.dir="/home/vmuser/Desktop/PHC7065SPR2021/WK4")
```

#Question 1

###Use the files "Patient_Drug_Admin_Date.csv" and "Patient_Drug390_Test_Flag.csv". We would like to know whether length of treatment and the distribution of HR/HER2 status varies based on the drug that the patient is receiving (whether they are receiving drug 390 or not). Write codes to compare: 1) the length of treatment by drug (drug 390 vs. other), and 2) the HR/HER status distribution by drug (drug 390 vs. other).
#### Assuming that the treatment length (unit: day) is determined by the length between the first admin_date and the last admin_date. For those with only one administer record, the length is therefore 0.
#### Note that patients with breast cancer can be categorized into 4 groups, based on their HER2 status and HR status (i.e. HR+ if either ER or PR is positive, HR- if both ER and PR are negative)

```{r}
library("RMySQL")

#import data
pdad<-read.csv("dat/Patient_Drug_Admin_Date.csv",sep = "\t", header=T, stringsAsFactors = F)
pdtf<-read.csv("dat/Patient_Drug390_Test_Flags.csv",sep = "\t", header=T, stringsAsFactors = F)

#convert chr to date
pdad$Drug_admin_date<-as.Date(pdad$Drug_admin_date,format = "%m/%d/%y")

#lower case colnames
colnames(pdad)<-tolower(colnames(pdad))
colnames(pdtf)<-tolower(colnames(pdtf))

#import data to mysql
pw <- "vmuser"
drv<-dbDriver("MySQL")
con<-dbConnect(drv,dbname="PHC7065DB",host="localhost",user="phc7065",password=pw)

dbWriteTable(con,"pdad",pdad,row.names=F,overwrite=T)
dbWriteTable(con,"pdtf",pdtf,row.names=F,overwrite=T)

#add hr+
q1<-"
ALTER TABLE pdtf
ADD hr SMALLINT;"

q2<-"
UPDATE pdtf 
  SET hr=
    CASE WHEN er_positive=1 OR pr_positive=1 THEN 1
       WHEN er_positive=0 AND pr_positive=0 THEN 0
       ELSE NULL
       END;
"
dbSendQuery(con,q1)
dbSendQuery(con,q2)

q<-"
SELECT *
FROM pdtf;
"
dbGetQuery(con,q)

```

```{r}
#Question 1) generate length of treatment
q<-"
SELECT c.drug_390_admin_flag, AVG(c.length) as meanlength
FROM
(SELECT b.patient_id, b.length, pdtf.drug_390_admin_flag
FROM
(SELECT a.patient_id, a.max_date-a.min_date as length
FROM
(SELECT pdad.patient_id, MAX(pdad.drug_admin_date) as max_date, MIN(pdad.drug_admin_date) as min_date
FROM pdad
GROUP BY patient_id) as a) as b
LEFT OUTER JOIN pdtf ON b.patient_id=pdtf.patient_id) as c
GROUP BY c.drug_390_admin_flag
;
"
dbGetQuery(con,q)
```

```{r}
#Question 2) HR/HER status by Drug
q<-"
SELECT drug_390_admin_flag, her2_positive, hr, COUNT(patient_id)
FROM pdtf
GROUP BY drug_390_admin_flag, her2_positive, hr
;
"
dbGetQuery(con,q)
```

#Question 2-1

###Use the files "Patient_Doctor_Product_Month.csv" and "Site.csv". Write codes that would generate a table showing the market share of each product by State (of the addresses of the cancer centers - sites) for the year of 2012

```{r}
#import data
pdpm<-read.csv("dat/Patient_Doctor_Product_Month.csv", header=T, stringsAsFactors = F)
site<-read.csv("dat/Site.csv", header=T, stringsAsFactors = F)

#lower case colnames
colnames(pdpm)<-tolower(colnames(pdpm))
colnames(site)<-tolower(colnames(site))

#import data to mysql
dbWriteTable(con,"pdpm",pdpm,row.names=F,overwrite=T)
dbWriteTable(con,"site",site,row.names=F,overwrite=T)

#Get the results
q<-"
SELECT a.site_state, a.product_id, SUM(a.quantity) as quantity
FROM
(SELECT pdpm.*, site.site_state
FROM pdpm
LEFT OUTER JOIN site ON pdpm.site_id=site.site_id) as a
WHERE a.month BETWEEN 201201 AND 201212
GROUP BY a.site_state, a.product_id
;
"
dbGetQuery(con,q)
```

#Question 2-2

###Continuing from question 2-1, what if there are NULL values in the State field and everything else remains the same (use the "Site2.csv" instead of "Site.csv")?

```{r}
#import data
site2<-read.csv("dat/Site2.csv", header=T, stringsAsFactors = F)

#lower case colnames
colnames(site2)<-tolower(colnames(site2))

#import data to mysql
dbWriteTable(con,"site2",site2,row.names=F,overwrite=T)

#Get the results
q<-"
SELECT a.site_state, a.product_id, SUM(a.quantity) as quantity
FROM
(SELECT pdpm.*, site2.site_state
FROM pdpm
LEFT OUTER JOIN site2 ON pdpm.site_id=site2.site_id) as a
WHERE a.month BETWEEN 201201 AND 201212
GROUP BY a.site_state, a.product_id
;
"
dbGetQuery(con,q)
```

#Question 2-3

###Continuing from question 2-1, there is a concern that some doctors use certain products in this market in off-label manners and you are told to keep strictly only to the Diagnosis code of 162.9 in this market share analysis. Please write the codes which would include the revised definition (use the files "Patient_Doctor_Product_Month.csv", "Patient_Doctor_Month_Diagnosis.csv", and "Site.csv")

```{r}
#import data
pdmd<-read.csv("dat/Patient_Doctor_Month_Diagnosis.csv", header=T, stringsAsFactors = F)

#lower case colnames
colnames(pdmd)<-tolower(colnames(pdmd))

#import data to mysql
dbWriteTable(con,"pdmd",pdmd,row.names=F,overwrite=T)

#get the results
q<-"
SELECT a.site_state, a.product_id, SUM(a.quantity) as quantity
FROM
(SELECT pdpm.*, site.site_state, pdmd.diagnosis_code
FROM pdpm
LEFT OUTER JOIN site ON pdpm.site_id=site.site_id
LEFT OUTER JOIN pdmd ON pdpm.patient_id=pdmd.patient_id AND pdpm.doctor_id=pdmd.doctor_id AND pdpm.month=pdmd.month) as a
WHERE (a.month BETWEEN 201201 AND 201212) AND (a.diagnosis_code=162.9)
GROUP BY a.site_state, a.product_id
;
"
dbGetQuery(con,q)
```

